\chapter{Theoretical Grounding}
\label{ch:TheoreticalGrounding}
This chapter introduces and explains terms and concepts relevant for this thesis. In order to follow the empirical research throughout the thesis, it is important to gain fundamental knowledge in the domains of financial markets, technical analysis, and machine learning.

\section{Financial Markets}
\label{sec:FinMarket}
Financial markets comprise of various concepts, terms and instruments. This section delves into the these aspects and elaborates on them.

\input{./parts/02_related_work/finma.tex}

\newpage
\section{Technical Analysis}
\label{sec:TechnicalAnalysis}
Technical analysis covers a wide range of concepts, techniques and tools that help traders and investors analyse financial markets and make informed trading decisions based on historical price, volume and other related market data. In order to follow the thesis, it is important to gain a basic knowledge of technical indicators from the field of technical analysis.

\input{./parts/02_related_work/ta.tex}

\newpage
\section{Algorithmic Trading}
\label{sec:AlgorithmicTrading}
There is no single, universally accepted definition of algorithmic trading. At its most basic level, it refers to the trading of financial instruments based on some formal algorithm \citep{hilpisch2020python}. An algorithm is a sequence of instructions that a computer must perform to solve a well-defined problem.

\input{./parts/02_related_work/algo_trading.tex}

\newpage
\section{Machine Learning}
\label{sec:MachineLearning}
This section provides an introduction to the most essential concept of this thesis, machine learning algorithms.

\input{./parts/02_related_work/dt.tex}

\input{./parts/02_related_work/gb.tex}

\input{./parts/02_related_work/xgboost.tex}

\input{./parts/02_related_work/ann.tex}
    
\input{./parts/02_related_work/rnn.tex}

\input{./parts/02_related_work/lstm.tex}

\input{./parts/02_related_work/af.tex}